---
title: "Tidy your Data: Join Data Sets"
description: |
  **Session 10**: Join Data Sets
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
<style>
  .blackbox {
    padding: 1em;
    background: #b0c4cf;
    color: black;
    border: 2px solid #455a64;
    border-radius: 7px;
  }
</style>

## 10.1 Tidying data

Often observations are spread across multiple data sets, and you will have to join data sets into one single table. 

:::: {.blackbox data-latex=""}
**Go through the RStudio Primer [Join Data Sets](https://rstudio.cloud/learn/primers/4.3)** and learn about the different functions to join data sets.
::::
<p>&nbsp;</p>

### Assignment

For this assignment, we will be joining the dataset **taxation** we used in Session 2 and 4 and a dataset from **Open Data Basel City**.
You can download this new dataset from [this webpage](https://data.bs.ch/explore/dataset/100011/information/?disjunctive.publikationsjahr&disjunctive.wohnviertel_id&disjunctive.wohnviertel_name&sort=-publikationsjahr) and call it **kennzahlen_basel**.

  1. Look at the two datasets. Join the **kennzahlen_basel** to the **taxation** dataset by year and quarter, retaining only the years that are present in both datasets (2015 to 2017).
  Your output should look like this:

``` {r echo = F, results = 'asis'}
library(tidyverse)
library(knitr)
library(kableExtra)

taxation <- read_csv("data/taxation.csv")
kennzahlen_basel <- read_csv2("data/kennzahlen_basel.csv") 

# change the variable name in kennzahlen_basel to match taxation
kennzahlen_basel <- kennzahlen_basel %>%
  rename(year = Publikationsjahr,
         quarter = `Wohnviertel Name`)

# join datasets by year and county
combined_data <- taxation %>%
  inner_join(kennzahlen_basel, by = c("year", "quarter"))

combined_data %>%
  kable() %>%
  kable_styling() %>%
  scroll_box(width = "700px", height = "300px")


```

``` {r echo = F, results = 'hide'}
### Solution
taxation <- read_csv("data/taxation.csv")
kennzahlen_basel <- read_csv2("data/kennzahlen_basel.csv") 

# change the variable name in kennzahlen_basel to match taxation
kennzahlen_basel <- kennzahlen_basel %>%
  rename(year = Publikationsjahr,
         quarter = `Wohnviertel Name`)

# join datasets by year and county
combined_data <- taxation %>%
  inner_join(kennzahlen_basel, by = c("year", "quarter"))
```






