---
title: "Work with Data: Derive Information with dplyr"
description: |
  **Session 5**: Derive Information with dplyr
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
<style>
  .blackbox {
    padding: 1em;
    background: #b0c4cf;
    color: black;
    border: 2px solid #455a64;
    border-radius: 7px;
  }
</style>

## 5.1 Derive Information with dplyr

In this session, you will continue learning data manipulation with `dplyr`.

:::: {.blackbox data-latex=""}
**Go through the RStudio Primer on [Derive Information with dplyr](https://rstudio.cloud/learn/primers/2.3),** and **complete the assignments below.**
::::
<p>&nbsp;</p>

### Assignment
 
Using the **taxation** dataset, complete the following tasks:

  1. Create a new column called `income_labels` from `income_mean`, where you have 4 new income categories **below 50000, 5000-79999, 8000-109999, and above 110000.** You can code the categories as numbers from 1 to 4. (*hint*: use `case_when()` in `mutate()` when you make the new columns.)
  
  2. How many quarters are there for each of the four categories by year? Use `group_by()` and `summarize()` to answer. You can make a table like this:
  
```{r echo = F, results = 'asis'}
library(tidyverse)
library(knitr)
library(kableExtra)
taxation <- read_csv("data/taxation.csv")
taxation = mutate(taxation, income_labels = case_when(income_mean < 50000 ~ "1",
                                              income_mean >= 50000 & income_mean < 80000 ~ "2",
                                              income_mean >= 80000 & income_mean < 110000 ~ "3",
                                              income_mean >= 110000  ~ "4"))
```
``` {r include = T}
new_taxation = taxation %>% 
  group_by(year, income_labels) %>%
  summarise(number_quarters = n())

new_taxation %>%
  kable() %>%
  kable_styling() %>%
  scroll_box(height = "300px")
  
```

```{r echo = F, results = 'hide'}
### Solution
taxation <- read_csv("data/taxation.csv")

# create new column
taxation = mutate(taxation, income_labels = case_when(income_mean < 50000 ~ "1",
                                              income_mean >= 50000 & income_mean < 80000 ~ "2",
                                              income_mean >= 80000 & income_mean < 110000 ~ "3",
                                              income_mean >= 110000  ~ "4"))

# count observations by year and quarter
new_taxation = taxation %>% 
  group_by(year, income_labels) %>%
  summarise(number_quarters = n()) %>%
  print()
```