---
title: "Tidy your Data: Separate and Unite Columns"
description: |
  **Session 9**: Separate and Unite Columns
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
<style>
  .blackbox {
    padding: 1em;
    background: #b0c4cf;
    color: black;
    border: 2px solid #455a64;
    border-radius: 7px;
  }
</style>

## 9.1 Tidying data

In the last session, you learned how to use `pivot_longer()` and `pivot_wider()` to tidy the layout of your data. In this session, you will learn how to separate or combine cells using `separate()` and `unite()`.

:::: {.blackbox data-latex=""}
**Go through the RStudio Primer [Separate and Unite Columns](https://rstudio.cloud/learn/primers/4.2).**
::::
<p>&nbsp;</p>

### Assignment

Reload the `billboard` dataset from the package `tidyr`.

  1. Separate the column `data.entered` into three new columns containing the corresponding values: `year`, `month` and `date`.
  2. Reunite the three columns back into `data.entered`, but leave the three columns in the dataset. 
  Your dataset should look like this:

```{r include=T}
library(tidyr)
library(knitr)
library(kableExtra)

billboard_new = billboard %>%
  separate(date.entered, into = c("year", "month", "date"), sep = "-")

billboard_new %>%
  unite(date.entered, year, month, date, sep = "-", remove = F) %>%
  kable() %>%
  kable_styling() %>%
  scroll_box(width = "700px", height = "300px")
``` 


```{r echo = F, results = 'hide'}
### Solution
billboard_new = billboard %>%
  separate(date.entered, into = c("year", "month", "date"), sep = "-")

billboard_new %>%
  unite(date.entered, year, month, date, sep = "-", remove = F)
```



