---
title: "Tidy your Data: Join Data Sets"
description: |
  **Session 10**: Join Data Sets
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## 10.1 Tidying data

Often observations are spread across multiple data sets, and you will have to join data sets into one single table. 

**Go through the RStudio Primer 4.3 [Separate and Unite Columns](https://rstudio.cloud/learn/primers/4.3)** and learn about the different functions to join data sets.

### Assignment

For this assignment, we will be joining the dataset **taxation** we used in Session 2 and 4 and a dataset from **Open Data Basel City**.
You can download this new dataset from [this webpage](https://data.bs.ch/explore/dataset/100011/information/?disjunctive.publikationsjahr&disjunctive.wohnviertel_id&disjunctive.wohnviertel_name&sort=-publikationsjahr) and call it **kennzahlen_basel**.

  1. Look at the two datasets. Join the **kennzahlen_basel** to the **taxation** dataset by year and quarter, retaining only the years that are present in both datasets (2015 to 2017).
  Your output should look like this:
```{r, include=FALSE}
options(width = 60)
local({
  hook_output <- knitr::knit_hooks$get('output')
  knitr::knit_hooks$set(output = function(x, options) {
    if (!is.null(options$max.height)) options$attr.output <- c(
      options$attr.output,
      sprintf('style="max-height: %s;"', options$max.height)
    )
    hook_output(x, options)
  })
})
```
``` {r include = T, echo = F, max.height='100px'}
library(tidyverse)

taxation <- read_csv("data/taxation.csv")
kennzahlen_basel <- read_csv2("data/kennzahlen_basel.csv") 

# change the variable name in kennzahlen_basel to match taxation
kennzahlen_basel <- kennzahlen_basel %>%
  rename(year = Publikationsjahr,
         quarter = `Wohnviertel Name`)

# join datasets by year and county
combined_data <- taxation %>%
  inner_join(kennzahlen_basel, by = c("year", "quarter")) %>%
  print()
```

### Solution
``` {r echo = T, results = 'hide'}
taxation <- read_csv("data/taxation.csv")
kennzahlen_basel <- read_csv2("data/kennzahlen_basel.csv") 

# change the variable name in kennzahlen_basel to match taxation
kennzahlen_basel <- kennzahlen_basel %>%
  rename(year = Publikationsjahr,
         quarter = `Wohnviertel Name`)

# join datasets by year and county
combined_data <- taxation %>%
  inner_join(kennzahlen_basel, by = c("year", "quarter"))
```






